---
title: "HW11: Education Level and Survey Weighting"
author: Elizabeth Polo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

# Introduction

When a poll or survey gets published, the results almost always include a note about sample size and methodology. One thing you'll often see: "results have been weighted to reflect the U.S. adult population." But what does weighting actually *do* to the numbers — and does it matter for your story?

In this assignment, you'll compare weighted and unweighted results from a 2020 survey on voting intentions, focusing on how education level relates to whether someone planned to vote.

## The Data

```{r load-data}
nonvoters_data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/nonvoters_data.csv")
glimpse(nonvoters_data)
```

Key variables:

- `weight`: Each respondent's survey weight. A weight of 1.0 means they count as exactly one person. A weight of 2.0 means their response counts double (they represent an undersampled group). A weight of 0.5 means their response counts half as much (they represent an oversampled group).
- `Q21`: Voting intention — 1 = Yes, 2 = No, 3 = Unsure
- `educ`: Education level — College, Some college, High school or less

## Task 1: Examine the Weights (4 points)

Before comparing weighted and unweighted results, let's understand what the weights look like.

```{r examine-weights}
summary(nonvoters_data$weight)

ggplot(nonvoters_data, aes(x = weight)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(
    title = "Distribution of Survey Weights",
    x = "Weight Value",
    y = "Number of Respondents"
  )
```

**Question 1:** Most weights cluster near 1.0, but some are much higher or lower. What does a weight of 0.2 tell you about that respondent's group in the sample? What about a weight of 3.5? Why might a pollster assign such extreme weights?

A weight of 0.2 in this sample tells me that this data is extremely oversampled, and 3.5 means that the respondant's group is extremely undersampled. This could be for specific demographics, like age, sexual orientation, fincances, etc. 

## Task 2: Education Distribution in the Sample (3 points)

```{r education-distribution}
nonvoters_data |>
  count(educ) |>
  mutate(percentage = round(n / sum(n) * 100, 1))
```

**Question 2:** College graduates are overrepresented in most surveys. Does that appear to be true here? Why might college graduates be more likely to respond to a survey than people with less education?

This does appear to be true. College graduates are at almost 40%, while the others are significantly lower. I think college grads are more likely to respond to surveys because they might've worked more with surveys during their schooling. Also, I feel like some surveys are more applicable to those with college degrees. Not too sure though, if I'm honest.

## Task 3: Unweighted Voting Intentions by Education (4 points)

This is what the results look like if you treat every respondent equally — no weights applied.

```{r unweighted}
unweighted_by_education <- nonvoters_data |>
  filter(!is.na(Q21), Q21 > 0, !is.na(educ)) |>
  group_by(educ, Q21) |>
  summarize(count = n(), .groups = "drop_last") |>
  mutate(total = sum(count),
         percentage = count / total * 100) |>
  ungroup()

unweighted_summary <- unweighted_by_education |>
  pivot_wider(
    id_cols = educ,
    names_from = Q21,
    values_from = percentage,
    names_prefix = "pct_"
  ) |>
  rename("Yes (%)" = pct_1, "No (%)" = pct_2, "Unsure (%)" = pct_3)

unweighted_summary
```

**Question 3:** Based only on these unweighted numbers, write a one-sentence headline for a story about education and voting intentions.

Survey suggests higher education leads to more voting.

## Task 4: Weighted Voting Intentions by Education (2 points)

Now apply the weights. Instead of counting respondents, we sum their weights — so someone with a weight of 2.0 contributes twice as much as someone with a weight of 1.0.

```{r weighted}
weighted_by_education <- nonvoters_data |>
  filter(!is.na(Q21), Q21 > 0, !is.na(educ)) |>
  group_by(educ, Q21) |>
  summarize(weighted_count = sum(weight), .groups = "drop_last") |>
  mutate(weighted_total = sum(weighted_count),
         weighted_percentage = weighted_count / weighted_total * 100) |>
  ungroup()

weighted_summary <- weighted_by_education |>
  pivot_wider(
    id_cols = educ,
    names_from = Q21,
    values_from = weighted_percentage,
    names_prefix = "pct_"
  ) |>
  rename("Yes (%)" = pct_1, "No (%)" = pct_2, "Unsure (%)" = pct_3)

weighted_summary
```

## Task 5: Visualize the Difference (4 points)

```{r visualization}
viz_data <- bind_rows(
  unweighted_by_education |>
    filter(Q21 == 1) |>
    mutate(Type = "Unweighted") |>
    select(Type, educ, percentage),
  weighted_by_education |>
    filter(Q21 == 1) |>
    mutate(Type = "Weighted",
           percentage = weighted_percentage) |>
    select(Type, educ, percentage)
)

ggplot(viz_data, aes(x = educ, y = percentage, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            position = position_dodge(width = 0.9), vjust = -0.5, size = 3) +
  labs(
    title = "Weighted vs. Unweighted: Share Planning to Vote, by Education",
    x = "Education Level",
    y = "% Saying Yes"
  ) +
  scale_fill_manual(values = c("Unweighted" = "#2166ac", "Weighted" = "#d6604d")) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Question 4:** Which education group shows the biggest shift between weighted and unweighted results? In percentage points, how large is that shift? Does weighting change the story, or does it just change the exact numbers?

The biggest shift was high school or less with a shift of 3.3%. Weighting doesn't change the story here. It just changes the numbers to be more accurate to the populaiton.

## Task 6: The UMD Belonging Survey (8 points)

Earlier in class this week, we looked at the 2024 UMD Belonging and Community Survey. That survey had a 23% response rate and the report acknowledged "statistically significant differences" between the sample and the UMD population — but applied no weighting to correct for them.

**Question 5:** Based on what you've learned in this assignment, write 2-3 paragraphs as if you were a reporter evaluating the UMD survey before writing a story based on it. Address: (1) what the lack of weighting means for how you'd use the results, (2) which groups in the UMD survey appeared most underrepresented based on Figures 1 and 2, and (3) what question you would ask the survey's authors before publishing.

This lack of weighting for such a big campus will poorly represent the total population of students. In their survey, especially under the racial and gender & sexuality sections, the university decided to leave out certain populations, like Native Americans, and underrepresented members of the LGBTQ community by comining them into "other" sections. 

By combining these minority groups into one, it doesn't actually cast the whole range of what makes UMD, UMD. It's hard to look at the survey and see it as accurate towards the population as it need to be. I would ask the publishers, why they decided to combine these members of our community into one, when the point of the survey is to capture belonging on campus. It's ironic because it just erases these groups in a way.
