---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: Elizabeth Polo
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

In the class activity, you tested a school board member's claim using data from Montgomery County, Pa. — and found a model that didn't hold up. Now you'll run the same analysis using data from Montgomery County, Md. Your job is to evaluate whether the Maryland data tells a different story, and what that means for how you'd report on the claim.

### Loading the Data

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```

### Task 1: Examine the Data First (2 points)

Before you plot anything, look at the data — you know why by now.

```{r}
summary(schools$`FARMS Pct`)
schools |> arrange(desc(`FARMS Pct`)) |> head(10)
```

Are there any data quality issues like the ones you found in the Pennsylvania data? How does the distribution of FARMS percentages compare between the two counties? Note anything that concerns you or gives you confidence in the data.

It seems to me that the schools with the highest FARMS percentage still has low math scores. The top 10 schools with the highest FARMS percentages are all under 40%. THe schools with lower FARMS percentages have significantly higher math scores.


### Task 2: Create a Scatterplot and Compare (5 points)

Replace the REPLACE_ME value with the correct variable name to create a scatterplot of FARMS percentage vs. math scores for Montgomery County, Md.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations",
       title = "Math Scores vs. FARMS Percentage, Montgomery County MD")
```

Describe what you see. Then compare this scatterplot specifically to the Pennsylvania one: what differences do you notice in the direction, strength, and spread of the relationship? The two counties share a name and are both suburban — what might explain why they look so different?

This scatterplot shows a negative correlation. It matches my theory that the lower the FARMS percentage, the higher the math scores. This plot has a much more stronger correlation to the Montgomery PA scatterplot. The other one was all over the place. There was a huge jump in numbers, and they had crazy outliers with schools in the 100% FARMS range. This is more legible. 


### Task 3: Linear Regression Model (5 points)

Replace the REPLACE_ME value with the correct variable name to run a linear regression model predicting math scores from FARMS percentage.

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)
summary(model)
```

Report the three key statistics: the p-value, the Adjusted R-squared, and the coefficient for FARMS Pct. Then interpret them together — not separately. What does the combination of these three numbers tell you about the relationship between poverty and math scores in Montgomery County, Md.? Is this model worth using? How does it compare to the Pennsylvania model?

The P-value is < 0.00000000000000022, the Adjusted R-squared is 0.6734 and the coefficient is -0.59511. All of these numbers tell me that there is a strong statistical relationship with my data. The P value is practically 0, which means that this isn't really happening due to chance. The R-squared is telling me that there's 67% accounted for, and there's about 33% unaccounted for. The coefficient says its in a negative correlation, which is exactly what is happening. THe other data for Montgomery, PA has not statistical significance at all.


### Task 4: Identifying Schools with the Biggest Residuals (6 points)

Replace the REPLACE_ME values with the correct variable names to calculate predicted values, residuals, and absolute residuals for each school. Then arrange the schools in descending order of absolute residuals.

```{r}
schools <- schools |>
  mutate(
    predicted = predict(model),
    residual = `% Met or Exceeded Math Expectations` - predicted,
    abs_residual = abs(residual)
  ) |>
  arrange(desc(abs_residual))

head(schools |> select(1, `FARMS Pct`, `% Met or Exceeded Math Expectations`, predicted, residual), 10)
```

Identify the school with the largest positive residual (doing much better than the model predicts) and the school with the largest negative residual (doing much worse). For each school, find its zip code and look it up on [CensusReporter.org](https://censusreporter.org).

For each of the two schools, answer:
- What does the demographic and economic profile of the zip code look like?
- Does anything in that profile help explain why the school is over- or under-performing the model?
- What would your first reporting question be to the principal of that school?

Largest positive residual:Fox Chapel Elementary, zip code - 20876
Largest negative residual:Germantown Elementary, zip code - 20874
The biggest differnece that stuck out to me was the median income. Germantown Elementary households are making six figures, while the median for Fox Chapel is in the 70,000s. Also, Germantown is more diverse with their residents. At least 50% of Germantown residents have a bachelors degree, while those in Fox Chapel are at around 28%. My first question for the principals would be what they think are the affecting the students the most. Is it finances? Family life? Asking the same question will give me greater insight to compare the schools.


### Task 5: Evaluate the Model and Write the Pitch (7 points)

You've now run this analysis on two counties. Return to the school board member's claim: that elementary schools with fewer students on free and reduced-price meals do better on standardized math tests.

Write a short story pitch of 4-6 sentences that addresses all of the following: Does the Maryland data support or contradict the claim? How does it compare to what you found in Pennsylvania? What would you ask the school board member if you called them for comment? And what's the most interesting specific finding from your residuals analysis: which school would you want to visit first, and why?

The school board member is right. Students not on the reduced price meals do do better on tests. I want to cover this story and see the discrepancies and compare it to the claims in Montgomery, PA. I will visit Carderock Springs Elementary first because they scored the highest on the exams.
