---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

ANSWER HERE
Just at a glance, the categories that seem the most newsworthy are theft, robbery other, and B & E residential. They have the biggest jumps in crime for specific months. In March to April, B & E residential went from 27-52. In August to September, robbery other went from 46-70. Finally, theft rose from 281-363 in September to October. These jumps catch my eye which makes finding stories easier.
### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
total_incidents <- pg_crime |> mutate(Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec)

average_incidents <- pg_crime |> mutate ((Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec)/12)

pg_crime_averages <- average_incidents
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

ANSWER HERE
Out of all the categories listed, I saw that accidents, auto stolen and theft from auto were crimes with the top three highest averages per month. The lowest categories are homicide, B&E other and sex offense. Robbery residential had no average that said N/A. From the top crimes in PG County, this data set tells me that most of the crimes committed are related to cars. I am curious as to why robbery residential has no average.

### Task 6: Testing a Theory
```{r}

```

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
june_change <- |> pg_crime_averages mutate(pct_change = (Jun - average_incidents)/ average_incidents * 100)
july_change <- |> pg_crime_averages mutate(pct_change = (Jul - average_incidents)/average_incidents * 100)
august_change <- |> pg_crime_averages mutate(pct_change = Aug - Average_incidents)/ average_incidents *100)
summer_crime_averages <- (june_change, july_chage, august_change)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

ANSWER HERE
I havent been able to solve the previous question, but from what I can gather from looking at the previous data sets, it seems like it has a plateau in numbers. It looks like it levels out.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE
New data reveals that the top crimes in Prince George's County are auto related. Within the past year, the highest recorded crimes include accidents, stolen auto and theft from auto. 
### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
