---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |>
  mutate(violent_rate_2019 = (violent2019 / pop2019) *10000, 
         property_rate_2019 = (property2019 / pop2019) *10000, 
         violent_rate2020 = (violent2020 / pop2020) *10000, 
         property_rate_2020 = (property2020 / pop2020) *10000)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

In the new columns, in 2019, Baltimore was the city with the most violent crime rate. In all other categories, Ocean City tops the charts. This is surprising because whenever I think of dangerious cities in Maryland, people always talk about Baltimore, not the state's vacation destination. I suppose the story here could be what about Ocean City makes it such a valuable place for crime.


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> 
  mutate(violent_change = ((violent_rate2020 - violent_rate_2019) / violent_rate_2019) * 100,
         property_change = ((property_rate_2020 - property_rate_2019) / property_rate_2019) * 100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

Crisfield had a 320% change, which was the highest percent change in the data. The lowest was Brookeville with a -100% change. I would write a story about Crisfield and why it had such an extreme switch in percent.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

This Maryland vacation spot might not be paradise after all.

New data shows that Ocean City, Maryland took the spot for the city with the most property crime for 2019 and 2020 as well as most violent crime in 2020. In just one year alone, violent crime for the city rose by more than 80%.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: Homework 5
```{r}
md_cities_percentiles <- md_cities_change |> mutate(violent_percentile = ntile(violent_rate2020, 100))

md_cities_percentiles |>
  summarize(IQR = IQR(violent_rate2020))
```
These results don't necessarily change my original story. Ocean City is still in the 100th percentile, which just gives me more data to add to my story.
