---
title: "HW8_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv

```{r}
M_and_M <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```

### Task 3: Sample 5 rows of data from the dataset
```{r}
sampled_candy <- population |>
  slice_sample(n = 1000)

sampled_candy
```


### Task 4: Calculate the mean age and standard deviation of the sample, using either the blue or white number of M&Ms
```{r}
sample_stats <- sampled_candy |>
  summarise(
    mean_age = mean(age),
    sd_age = sd(age)
  )

sample_stats
```


### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4
```{r}
candy_population_stats <- population |>
  summarise(
    mean_age = mean(age),
    sd_age = sd(age)
  )

candy_population_stats
```


### Task 6: Compare the sample and population statistics

Compare the two sets of means and standard deviations. Are they similar?

The two means and deviations are the exact same for me.

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

The data was the exactly the same for the total population. I think this is because the sample size is really small. Only 5. This tells me that if there were a larger sample size, then it would represent the populaiton more.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
