---
title: "final_project"
output: html_document
---

Here's a link to all the crime pages for each Big 10 university I found:
"https://docs.google.com/document/d/18XJh4Ae4NbCxPv01ecx6DyTsmicGlgGASVVTIDo0FB8/edit?tab=t.0"

Notes: Each university has different means of documenting public indecency/voyeurism. Some don't even identify it at all. This is def causing some trouble. 

I could shift points, and look at instances of other forms of sexual assault, like rape, fondling, stalking or domestic violence. Clery Act lists all these crimes as sexual assault categories.

Options: focus on one sexual assault categories, or focus on all sexual assault categories.

Update 4/19/26:

made excel sheet with data going back to 2001. 
```{r}
library(tidyverse)
```
Criminal_Offenses_On_campus_2_405_final <- read_csv("project/Criminal_Offenses_On_campus (2) 405 final.csv")

Hypothesis - How much did sexual assault reports show up after the 2017 Me Too movement?

```{r}
Criminal_offenses <- read_csv("https://raw.githubusercontent.com/Elizabeth-ops-ai/jour405_spring26/refs/heads/main/project/Criminal_Offenses_On_campus%20(2)%20405%20final.csv")
```
Now, I'm going to explore my data:

```{r}
# Get a summary of the data
glimpse(Criminal_offenses)

# Check for any missing values
colSums(is.na(Criminal_offenses))

# Check the total assault range in our data
min_total <- min(Criminal_offenses$`Total Assault`)
max_total <- max(Criminal_offenses$`Total Assault`)
cat("Data ranges from", min_total, "to", max_total)

```
It seems like 2023-02-06 UTC is the max date, and 2022-02-06 UTC is the min date.

Next, I want to figure out the correlation between the year, institution and total sexual assault.

```{r}
Criminal_offenses |> summarize(correlation = cor(`Survey year`, `Total Assault`, method="pearson"))
```

The correlation is 0.16. Not too strong of a correlation at all. 

I still want to see the data on a scatterplot though. I'll make that now.

```{r}
Criminal_offenses |>
  ggplot(aes(x = `Survey year`, y = `Total Assault`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Survey Year", y = "Total Assault",
       title = "Survey years vs. Total assault, BIG10 Schools")
```

There looks like a slight positive correlation. After 2017, the year of the Me Too Movement, there looks like some crazy spikes in outliers. Let's take a look at that.

```{r}
Filtered_offenses <- Criminal_offenses |> filter(`Total Assault` > 250)

model_filtered <- lm(`Total Assault` ~ `Survey year`, data = Filtered_offenses)
summary(Filtered_offenses)
```
```{r}
cat("Full dataset Criminal_offenses:\n")
cat("  Coefficient:", coef(model)[2], "\n")
cat("  p-value:", summary(Criminal_offenses)$coefficients[2,4], "\n")
cat("  Adjusted R-squared:", summary(Criminal_offenses)$adj.r.squared, "\n\n")

cat("Filtered dataset model (Total Assault < 250):\n")
cat("  Coefficient:", coef(model)[2], "\n")
cat("  p-value:", summary(model)$coefficients[2,4], "\n")
cat("  Adjusted R-squared:", summary(model)$adj.r.squared, "\n")
```

I don't understand the error here.

Let's also run a pre and post 2017 test.
```{r}
pre <- Criminal_offenses |> filter(`Survey year` < 2017) |> pull(`Total Assault`)
post <- Criminal_offenses |> filter(`Survey year` >= 2017) |> pull(`Total Assault`)

t.test(pre, post)
```
In the case of this data here with the p-value, there's not a lot of significance at all. 

The seven outliers are all schools that boomed following the Me Too movement in 2017. It ranged from from 6 years from 2018-2021 and 2023. Ohio and Michigan are the two largest numbers of assault numbering in the thousands. After some research, these two schools employeed doctors who assaulted and raped their patients: Dr. Richard Strauss, a team doctor who worked at Ohio from 1978 to 1998, and former sports Dr. Robert Anderson, who worked at Michigan from 1966-2003.
