---
title: "final_project"
output: html_document
---

Here's a link to all the crime pages for each Big 10 university I found:
"https://docs.google.com/document/d/18XJh4Ae4NbCxPv01ecx6DyTsmicGlgGASVVTIDo0FB8/edit?tab=t.0"

Notes: Each university has different means of documenting public indecency/voyeurism. Some don't even identify it at all. This is def causing some trouble. 

I could shift points, and look at instances of other forms of sexual assault, like rape, fondling, stalking or domestic violence. Clery Act lists all these crimes as sexual assault categories.

Options: focus on one sexual assault categories, or focus on all sexual assault categories.

Update 4/19/26:

made excel sheet with data going back to 2001. 
```{r}
library(tidyverse)
```
Criminal_Offenses_On_campus_2_405_final <- read_csv("project/Criminal_Offenses_On_campus (2) 405 final.csv")

Hypothesis - How much did sexual assault reports show up after the 2017 Me Too movement?

```{r}
Criminal_offenses <- read_csv("https://github.com/Elizabeth-ops-ai/jour405_spring26/commit/4de541fa8dadc460c2e0f1905cd326f384e8a24d")
```

